---
title: "HW8_G11_CameronMiller"
author: "Cameron Miller"
date: "2022-11-01"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA)
library(dplyr)
library(wooldridge)
library(stargazer)
```

\newpage

## Question 1
Because funds and risktol are positively correlated, we can assume that 
the asymptotic bias in $\hat{\beta}_1$ is positive.

If we dropped risktol, then $\beta_1$ will become biased

## Question 2

### Part I

```{r}
df <- wage1
reg <- lm(wage ~ educ + exper + tenure, df)
stargazer(reg, type = "text")

my_resid <- reg$residuals

hist(my_resid, c= 'yellow', main = "Residual Model For Level-Level")

```


### Part II

```{r}

reg_log <- lm(log(wage) ~ educ + exper + tenure, df)
stargazer(reg_log, type = "text")
my_resid_log <- reg_log$residuals

hist(my_resid_log, c= 'light green', main = "Residual Model For Log-Level")
```


### Part III
Assumption number 6 is closer to being satisfied with the log-level model. 
This is because it is more normally distributed than the level- level model. 

## Question 3


### Part I

```{r}
range(econmath$score)
```
Logically, because it is a percentage, the low / high values can be 0 / 100.
By taking the range, we can see the actual low / high values are 19.5 / 98.4

### Part II

```{r}
df2 <- econmath
reg_econ <- lm(score ~ colgpa + actmth + acteng, df2)
hist(reg_econ$residuals)
```
It is not normally distributed and therefor cannot hold under MLR.6
If assumption 6 is not holding, there is no t-distribution and therefor you 
cannot talk about significance. 

### Part III

```{r}
summary(reg_econ)
dim(econmath)
```

The t-stat for acteng is .466 and the p value is 64%.

You can defend by saying that there are 856 observations and the central
limit theorem should be present, so the numbers listed can be trusted. 











